\bf Faster Neural Machine Translation Inference

### 2018-01-29 fast-beam/fast-beam.tex
fast beam paper

Firstly, the number of classes in many deep-learning applications is small. However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language. For example, ~\cite{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 subword units. This makes the output layer of NMT models very computationally expensive. ??? shows the breakdown of amount of time during translation our NMT system; over 40\% of the time is involved in calculating the activation, softmax and the beam search. We will look at optimizations which explicitly target the output layer.

@@

We shall leave the matrix multiplication for the next proposal and future work, and concentrate on the last three steps, the outline of which are shown in Figures~\ref{algo:Add Bias Term} to ~\ref{algo:Find best}.

@@

As can be seen, the operations iterate over the matrix p five times - once to add the bias, three times to calculate the softmax, and once to search for the best class. We shall fused the three functions and combine the matrix access into two iterations, reducing memory bandwidth. Kernel fusion~\citep{Guevara2009EnablingTP} is a well practised optimization technique.

Secondly, we make use of the fact that softmax and exp are monotonic functions therefore we can avoid the search for the argmax in Figure~\ref{algo:Find best} as it will be the same as that in Figure~\ref{algo:Calculate softmax}.

@@

\begin{figure} [h]
\begin{algorithmic}
\REQUIRE activation vector $p$, bias vector $b$

>>>>

Firstly, the number of classes in many deep-learning applications is small. However, the number of classes in NMT models is typically in the tens or hundreds of thousands, corresponding to the vocabulary size of the output language. For example, ~\cite{sennrich-haddow-birch:2016:P16-12} experimented with target vocabulary sizes of 60,000 and 90,000 subword units. This makes the output layer of NMT models very computationally expensive. Figure~\ref{fig:pie-time} shows the breakdown of amount of time during translation our NMT system; nearly 70\% of the time is involved in the output layer. We will look at optimizations which explicitly target the output layer.
\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{pie-time.png}} 
\end{tabular}
\caption{Time spent during translation (beam size 5)}
\label{fig:pie-time}
\end{figure}

@@

We shall leave step 1 for future work and focus on the last three steps, the outline for which are shown in Figures~\ref{algo:Add Bias Term} to ~\ref{algo:Find best}.

@@

As can be seen, the operations iterate over the matrix p five times - once to add the bias, three times to calculate the softmax, and once to search for the best class. We shall use a popular method, kernel fusion~\citep{Guevara2009EnablingTP}, to optimize the output layer.

Secondly, we make use of the fact that softmax and $\exp$ are monotonic functions therefore we can move the search for the best class from Figure~\ref{algo:Find best} to Figure~\ref{algo:Calculate softmax} while $max$ is being sought.

@@

\begin{figure} [h]
\begin{algorithmic}
\REQUIRE vector $p$, bias vector $b$
